---
title: "Medulloblastoma — Network (Late Integration)"
subtitle: "SCPCS000136 · SCPCS000139 · GENIE3 single-cell inference · Consensus by intersection"
author: |
  **Francesco Cecere**
  *Senior Bioinformatic Researcher*
  AORN Santobono-Pausilipon
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = FALSE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center",
  dpi       = 96
)
```

```{r libraries, include=FALSE}
library(Seurat)
library(SingleCellExperiment)
library(scGraphVerse)
library(igraph)
library(SummarizedExperiment)
library(tidyverse)
library(ggplot2)
library(ggraph)
library(tidygraph)
library(ggrepel)
library(RColorBrewer)
library(viridis)
library(visNetwork)
library(DT)
library(plotly)
library(patchwork)
library(downloadthis)
library(org.Hs.eg.db)
library(AnnotationDbi)

set.seed(42)

# ── Paths ──────────────────────────────────────────────────────────────────────
# Set base_dir to the folder containing your Seurat RDS files.
# Using "." means the current working directory (where this .Rmd lives).
base_dir  <- "."
net_dir   <- file.path(base_dir, "networks")
plots_dir <- file.path(base_dir, "plots")

dir.create(net_dir,   recursive = TRUE, showWarnings = FALSE)
dir.create(plots_dir, recursive = TRUE, showWarnings = FALSE)

# ── Parameters ────────────────────────────────────────────────────────────────
n_genes  <- 3000   # top expressed genes selected by selgene()
n_cores  <- 15     # GENIE3 threads — adjust to your machine

# ── Colour scheme ─────────────────────────────────────────────────────────────
sample_colors <- c(
  "SCPCS000136" = "#2196F3",
  "SCPCS000139" = "#FF9800"
)

# ── Helpers ───────────────────────────────────────────────────────────────────
dl_gg <- function(p, fname, w = 10, h = 7, label = "Download PNG (300 dpi)") {
  path <- file.path(plots_dir, paste0(fname, ".png"))
  suppressMessages(ggsave(path, plot = p, dpi = 300, width = w, height = h, bg = "white"))
  downloadthis::download_file(path = path, output_name = fname,
    button_label = label, button_type = "primary",
    has_icon = TRUE, icon = "fa fa-download")
}

dt_table <- function(df, caption = NULL, pageLength = 15) {
  DT::datatable(df, caption = caption, extensions = "Buttons",
    filter = "top", rownames = FALSE,
    options = list(dom = "Bfrtip", buttons = list("csv", "excel"),
                   pageLength = pageLength, scrollX = TRUE, autoWidth = TRUE),
    class = "compact hover stripe")
}

fmt_time <- function(secs) {
  if (secs < 60) sprintf("%.1f seconds", secs)
  else           sprintf("%.1f minutes", secs / 60)
}
```

---

# 1. Load Data

> **Prerequisites:** This script expects two Seurat RDS files in `base_dir`:
>
> - `seurat_SCPCS000136.rds`
> - `seurat_SCPCS000139.rds`
>
> These are produced by `create_seurat_objects.Rmd` starting from the processed
> single-cell data available on the
> [Childhood Cancer Data Lab (SCPCA) portal](https://scpca.alexslemonade.org/)
> for samples **SCPCS000136** and **SCPCS000139** (project SCPCP000009,
> Medulloblastoma, WHO grade 4). Raw count matrices are used throughout —
> no normalisation is applied before network inference.

The two per-sample Seurat objects are loaded here. No cell-type subsetting is applied — all cells from each sample are passed directly to the network inference engine. The raw count matrix is extracted from the `RNA` assay of each Seurat object.

```{r load-data}
seu_136 <- readRDS(file.path(base_dir, "seurat_SCPCS000136.rds"))
seu_139 <- readRDS(file.path(base_dir, "seurat_SCPCS000139.rds"))

mat_136 <- GetAssayData(JoinLayers(seu_136, assay = "RNA"), assay = "RNA", layer = "counts")
mat_139 <- GetAssayData(JoinLayers(seu_139, assay = "RNA"), assay = "RNA", layer = "counts")

cat(sprintf("SCPCS000136 : %d cells,  %d genes\n", ncol(mat_136), nrow(mat_136)))
cat(sprintf("SCPCS000139 : %d cells,  %d genes\n", ncol(mat_139), nrow(mat_139)))
```

---

# 2. Gene Selection with `selgene()`

Rather than using highly variable genes, we select the **top `r n_genes` expressed genes** using `scGraphVerse::selgene()`. Gene selection is performed on the combined count matrix (both samples pooled), so both networks are inferred on an identical gene universe — a prerequisite for a meaningful intersection-based consensus. Mitochondrial and ribosomal genes and MALAT1 are excluded. Gene identifiers are converted from Ensembl IDs to HGNC symbols before selection so that `selgene()` can correctly filter MT/ribosomal genes by name prefix and downstream KEGG enrichment receives gene symbols.

```{r gene-selection}
# Convert ENSG IDs → gene symbols on the shared gene universe
all_genes <- intersect(rownames(mat_136), rownames(mat_139))
sym_map <- mapIds(org.Hs.eg.db, keys = all_genes,
                  column = "SYMBOL", keytype = "ENSEMBL", multiVals = "first")
sym_map <- sym_map[!is.na(sym_map)]

mat_136_sym <- mat_136[names(sym_map), , drop = FALSE]
mat_139_sym <- mat_139[names(sym_map), , drop = FALSE]
rownames(mat_136_sym) <- sym_map
rownames(mat_139_sym) <- sym_map

# Gene selection on symbol-named combined matrix
mat_combined <- cbind(mat_136_sym, mat_139_sym)
mat_combined <- as.matrix(mat_combined)
mat_combined <- mat_combined[rownames(mat_combined) != "MALAT1", ]

selected_genes <- selgene(mat_combined, top = n_genes, remove_mt = TRUE, remove_rib = TRUE)
selected_genes <- intersect(selected_genes,
                             intersect(rownames(mat_136_sym), rownames(mat_139_sym)))

mat_136_sel <- mat_136_sym[selected_genes, , drop = FALSE]
mat_139_sel <- mat_139_sym[selected_genes, , drop = FALSE]

cat(sprintf("ENSG IDs with symbol mapping : %d\n", length(sym_map)))
cat(sprintf("Genes selected by selgene()  : %d\n", length(selected_genes)))
cat(sprintf("Final gene universe          : %d\n", nrow(mat_136_sel)))
```

---

# 3. Network Inference with GENIE3

Both samples are passed to the scGraphVerse pipeline **together as a named list** via `create_mae()`. `infer_networks()`, `generate_adjacency()`, `symmetrize()`, and `cutoff_adjacency()` each process the full list internally. A single `system.time()` call wraps the entire inference step so the wall-clock time for processing both samples simultaneously is captured.

```{r infer-networks}
mae <- create_mae(list(
  "SCPCS000136" = as.matrix(mat_136_sel),
  "SCPCS000139" = as.matrix(mat_139_sel)
))

set.seed(22172509)
time_inference <- system.time({
  networks <- infer_networks(
    count_matrices_list = mae,
    method              = "GENIE3",
    nCores              = n_cores
  )
})

elapsed_total <- time_inference[["elapsed"]]
cat(sprintf("Joint inference completed in %s\n", fmt_time(elapsed_total)))
```


```{r timing-table}
timing_df <- tibble(
  Sample       = c("SCPCS000136", "SCPCS000139", "Total (joint)"),
  Cells        = c(ncol(mat_136_sel), ncol(mat_139_sel), ncol(mat_136_sel) + ncol(mat_139_sel)),
  Genes        = nrow(mat_136_sel),
  Threads      = n_cores,
  `Time (s)`   = c(NA, NA, round(elapsed_total, 1)),
  `Time (min)` = c(NA, NA, round(elapsed_total / 60, 2))
)
dt_table(timing_df, caption = "GENIE3 inference runtime — both samples processed jointly")
```

The total joint inference time is split equally between the two samples as a per-matrix estimate. This avoids re-running GENIE3 and is a reasonable approximation given the similar cell counts. The per-sample times are saved to a CSV for import in the early integration report.

```{r timing-per-sample}
elapsed_per_sample <- elapsed_total / 2

timing_late <- tibble(
  sample   = c("SCPCS000136", "SCPCS000139"),
  label    = c("S136 (Late)", "S139 (Late)"),
  strategy = "Late",
  n_cells  = c(ncol(mat_136_sel), ncol(mat_139_sel)),
  n_genes  = nrow(mat_136_sel),
  time_s   = elapsed_per_sample,
  time_min = round(elapsed_per_sample / 60, 3)
)

cat(sprintf("Per-sample estimate (total / 2): %s\n", fmt_time(elapsed_per_sample)))
write_csv(timing_late, file.path(net_dir, "inference_timing_late.csv"))
cat("Saved: inference_timing_late.csv\n")
```

```{r adjacency-pipeline}
wadj_se   <- generate_adjacency(networks)
swadj_se  <- symmetrize(wadj_se, weight_function = "mean")
binary_se <- cutoff_adjacency(
  count_matrices     = mae,
  weighted_adjm_list = swadj_se,
  n                  = 5,
  method             = "GENIE3",
  quantile_threshold = 0.99,
  nCores             = n_cores
)
```

---

# 4. Consensus Network

`create_consensus()` is called with `method = "vote"` on the multi-assay binary result. With two samples in the list, the voting rule requires an edge to receive votes from **both** networks to be retained — equivalent to the strict intersection. Only edges present in both SCPCS000136 and SCPCS000139 survive, making the consensus a robust shared regulatory backbone.

```{r build-consensus}
consensus_se  <- create_consensus(binary_se, method = "vote")
consensus_mat <- assay(consensus_se)

consensus_g <- graph_from_adjacency_matrix(consensus_mat, mode = "undirected")
consensus_g <- simplify(consensus_g)
isolated    <- which(degree(consensus_g) == 0)
if (length(isolated) > 0) consensus_g <- delete_vertices(consensus_g, isolated)

V(consensus_g)$degree      <- degree(consensus_g)
V(consensus_g)$betweenness <- betweenness(consensus_g, normalized = TRUE)

cat("Assay names in binary_se:", paste(assayNames(binary_se), collapse = ", "), "\n")

A136 <- assay(binary_se, 1)
A139 <- assay(binary_se, 2)

cat(sprintf("Edges in S136    : %d\n", sum(A136) / 2))
cat(sprintf("Edges in S139    : %d\n", sum(A139) / 2))
cat(sprintf("Consensus (both) : %d\n", ecount(consensus_g)))
cat(sprintf("Consensus nodes  : %d\n", vcount(consensus_g)))

saveRDS(consensus_se, file.path(net_dir, "consensus_se.rds"))
saveRDS(consensus_g,  file.path(net_dir, "consensus_network.rds"))
```

## 4.1 Network Topology Statistics

```{r topology-stats}
deg_vals <- degree(consensus_g)

topo_df <- tibble(
  Metric = c(
    "Nodes (genes)", "Edges", "Average degree", "Max degree", "Density",
    "Average clustering coefficient", "Average shortest path length",
    "Number of components", "Giant component size"
  ),
  Value = c(
    vcount(consensus_g), ecount(consensus_g),
    round(mean(deg_vals), 2), max(deg_vals),
    round(graph.density(consensus_g), 5),
    round(transitivity(consensus_g, type = "average"), 4),
    round(mean_distance(consensus_g, directed = FALSE), 3),
    components(consensus_g)$no,
    max(components(consensus_g)$csize)
  )
)

dt_table(topo_df, caption = "Global topology of the consensus network", pageLength = 10)
```

## 4.2 Degree Distribution

```{r degree-distribution, fig.height=5}
deg_df <- tibble(gene = names(deg_vals), degree = as.integer(deg_vals))

p_deg <- ggplot(deg_df, aes(x = degree)) +
  geom_histogram(bins = 40, fill = "#2980b9", colour = "white", linewidth = 0.3) +
  scale_x_continuous(breaks = scales::pretty_breaks(8)) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank()) +
  labs(title = "Degree distribution — consensus network",
       x = "Degree (number of connections)", y = "Number of genes")

ggplotly(p_deg)
```

```{r degree-dist-dl}
dl_gg(p_deg, "consensus_degree_distribution", w = 9, h = 5)
```

---

# 5. Network Visualisation

## 5.1 Global Network Plot

The full consensus network is laid out using the Fruchterman–Reingold force-directed algorithm. Node size and colour encode degree: highly connected hub genes appear larger and in a darker shade. Only the top 30 hubs are labelled.

```{r network-plain, fig.height=10, fig.width=12}
set.seed(42)

tidy_g <- as_tbl_graph(consensus_g) %>%
  activate(nodes) %>%
  mutate(
    degree      = centrality_degree(),
    betweenness = centrality_betweenness(normalized = TRUE),
    is_hub      = degree >= quantile(degree, 0.95)
  )

top_hubs <- tidy_g %>%
  activate(nodes) %>%
  as_tibble() %>%
  slice_max(degree, n = 30) %>%
  pull(name)

p_net <- ggraph(tidy_g, layout = "fr") +
  geom_edge_link(colour = "#B0BEC5", alpha = 0.25, width = 0.35) +
  geom_node_point(aes(size = degree, colour = degree), alpha = 0.85) +
  scale_colour_viridis_c(option = "plasma", name = "Degree",
                          guide = guide_colourbar(barwidth = 8, barheight = 0.5,
                                                  title.position = "top")) +
  scale_size_continuous(range = c(1.5, 9), name = "Degree",
                         guide = guide_legend(override.aes = list(colour = "#2980b9"))) +
  theme_graph(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.box      = "horizontal",
    plot.title      = element_text(face = "bold", size = 14),
    plot.subtitle   = element_text(size = 10, colour = "#555")
  ) +
  labs(
    title    = "Medulloblastoma consensus co-expression network — Late integration",
    subtitle = sprintf("%d genes · %d edges", vcount(consensus_g), ecount(consensus_g))
  )

p_net
```

```{r network-plain-dl}
dl_gg(p_net, "consensus_network_plain", w = 14, h = 12)
```

---

# 6. Community Detection

Community detection is performed directly on the consensus `SummarizedExperiment` object using **`scGraphVerse::community_path()`**. The function applies the Louvain algorithm, assigns genes to communities, and immediately runs **KEGG pathway enrichment** (`clusterProfiler`) on each community with at least 5 genes.

```{r community-detection}
comm_result <- community_path(
  adj_matrix  = consensus_se,
  methods     = "louvain",
  pathway_db  = "KEGG",
  organism    = "human",
  genes_path  = 5,
  plot        = TRUE,
  verbose     = FALSE
)

builtin_plt <- recordPlot()
png(file.path(plots_dir, "consensus_community_builtin.png"),
    width = 12, height = 10, units = "in", res = 300, bg = "white")
replayPlot(builtin_plt)
invisible(dev.off())
cat("Saved: consensus_community_builtin.png\n")

memb        <- comm_result$communities$membership
best_method <- comm_result$communities$best_method

cat(sprintf("Algorithm   : %s\n", best_method))
cat(sprintf("Communities : %d\n", length(unique(memb))))
cat(sprintf("Modularity  : %.4f\n", modularity(comm_result$graph, memb)))

comm_sizes <- sort(table(memb), decreasing = TRUE)
comm_df <- tibble(
  Community_id = names(comm_sizes),
  `N genes`    = as.integer(comm_sizes)
) %>%
  mutate(
    Rank  = row_number(),
    Label = ifelse(Rank <= 3,
                   paste0("Community ", Rank, " (n=", `N genes`, ")"),
                   "Other")
  )

label_map <- setNames(comm_df$Label, comm_df$Community_id)

V(consensus_g)$community  <- as.character(memb[V(consensus_g)$name])
V(consensus_g)$comm_label <- label_map[V(consensus_g)$community]
V(consensus_g)$comm_label[is.na(V(consensus_g)$comm_label)] <- "Other"

dt_table(comm_df %>% dplyr::select(Rank, `Community ID` = Community_id, `N genes`, Label),
         caption = "Community sizes (scGraphVerse community_path — Louvain)")
```

## 6.1 Network Coloured by Community

```{r network-community, fig.height=11, fig.width=13}
n_top      <- min(3, sum(comm_df$Rank <= 3))
top_labels <- unique(comm_df$Label[comm_df$Rank <= 3])
comm_pal   <- setNames(
  RColorBrewer::brewer.pal(max(n_top, 3), "Set1")[seq_len(n_top)],
  top_labels
)

tidy_comm <- as_tbl_graph(consensus_g) %>%
  activate(nodes) %>%
  mutate(
    degree     = centrality_degree(),
    comm_label = V(consensus_g)$comm_label
  ) %>%
  filter(comm_label != "Other")

set.seed(42)
p_comm <- ggraph(tidy_comm, layout = "fr") +
  geom_edge_link(colour = "grey", alpha = 0.30, width = 0.35) +
  geom_node_point(aes(size = degree, colour = comm_label), alpha = 0.90) +
  scale_colour_manual(values = comm_pal, name = "Community",
                       guide = guide_legend(override.aes = list(size = 4), ncol = 2)) +
  scale_size_continuous(range = c(1.5, 9), name = "Degree",
                         guide = guide_legend(override.aes = list(colour = "#607D8B"))) +
  theme_graph(base_size = 12) +
  theme(
    legend.position = "right",
    plot.title      = element_text(face = "bold", size = 14),
    plot.subtitle   = element_text(size = 10, colour = "#555")
  ) +
  labs(
    title    = "Medulloblastoma consensus network — communities (Late integration)",
    subtitle = sprintf("%d communities · modularity = %.3f",
                       length(unique(memb)), modularity(comm_result$graph, memb))
  )

p_comm
```

```{r network-community-dl}
dl_gg(p_comm, "consensus_network_communities", w = 16, h = 12)
```

## 6.2 KEGG Pathway Enrichment per Community

`community_path()` runs pathway enrichment on every community with ≥ 5 genes. The table below collects all significant results (adjusted p < 0.05) across communities, ranked by gene ratio within each community.

```{r pathway-enrichment-table}
path_df <- imap_dfr(comm_result$pathways, function(res, comm_id) {
  if (is.null(res) || nrow(as.data.frame(res)) == 0) return(NULL)
  as.data.frame(res) %>%
    as_tibble() %>%
    mutate(Community = label_map[comm_id]) %>%
    dplyr::select(Community, ID, Description, GeneRatio, BgRatio,
           pvalue, p.adjust, qvalue, geneID, Count)
}) %>%
  filter(p.adjust < 0.05) %>%
  arrange(Community, p.adjust)

if (nrow(path_df) > 0) {
  dt_table(path_df %>% dplyr::select(-geneID),
           caption = "KEGG pathway enrichment — significant terms per community (adj. p < 0.05)")
} else {
  cat("No significant KEGG terms found at adj. p < 0.05.\n")
}
```

```{r pathway-enrichment-dotplot, fig.height=9, fig.width=13}
if (nrow(path_df) > 0) {
  p_pathway <- path_df %>%
    group_by(Community) %>%
    slice_min(p.adjust, n = 3, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(
      GeneRatio_num = map_dbl(GeneRatio, ~ eval(parse(text = .x))),
      Description   = str_trunc(Description, 45)
    ) %>%
    ggplot(aes(x = GeneRatio_num,
               y = reorder(Description, GeneRatio_num),
               colour = p.adjust, size = Count)) +
    geom_point() +
    facet_wrap(~ Community, scales = "free_y", ncol = 2) +
    scale_colour_gradient(low = "#E53935", high = "#90CAF9", name = "adj. p-value") +
    scale_size_continuous(name = "Gene count", range = c(2, 8)) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text       = element_text(face = "bold", size = 9),
      axis.text.y      = element_text(size = 8),
      panel.grid.minor = element_blank()
    ) +
    labs(title    = "KEGG pathway enrichment — top 3 terms per community (Late integration)",
         subtitle = "Dot size = gene count · Colour = adjusted p-value",
         x = "Gene Ratio", y = NULL)

  ggplotly(p_pathway)
  ggsave(file.path(plots_dir, "consensus_pathway_enrichment.png"),
         plot = p_pathway, dpi = 300, width = 14, height = 10, bg = "white")
  cat("Saved: consensus_pathway_enrichment.png\n")
} else {
  cat("No significant pathways to plot.\n")
}
```

## 6.3 Top Hub Genes per Community

```{r community-hubs-table}
hub_per_comm <- tidy_comm %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(comm_label != "Other") %>%
  group_by(comm_label) %>%
  slice_max(degree, n = 10, with_ties = FALSE) %>%
  ungroup() %>%
  dplyr::select(Community = comm_label, Gene = name, Degree = degree) %>%
  arrange(Community, desc(Degree))

dt_table(hub_per_comm, caption = "Top 10 hub genes per community")
```
